---
title: "region-scoping"
author: "jt-miller"
format: html
editor: visual
---

## Region Scoping

An overview of ecoregions & Neon sites of interest concerning the Pollen Database project

Libraries

```{r}
library(sf)
library(data.table)
library(tidyverse)
```

Bring in the Ecoregions & Neon Sites (Sites are depicted as centroids)

```{r}

na_ecoregions_3 <- sf::read_sf("/home/jt-miller/Pollen-Project/pollen-project-mapping/raw-data/Ecoregions/na-level3-ecoregions/NA_CEC_Eco_Level3.shp") %>% 
                      sf::st_transform(crs = 4326) # Transform into WGS84 from native projection (Lambert Azimuthal Equal Area)
neon_sites <- read.csv("/home/jt-miller/Pollen-Project/pollen-project-mapping/raw-data/NEON-sites/NEON_Field_Site_Metadata_20230309.csv")
```

Only Retain NEON sites of interest

```{r}
# We're focusing plants & insects, so we want sites that collect terrestrial data
unique(neon_sites$field_site_type)
terr_site_types <- c("Gradient Terrestrial", "Core Terrestrial")
neon_sites_t <- neon_sites %>% 
  filter(field_site_type %in% terr_site_types)
```

Read in as Spatial Data

```{r}
neon_sites_sp <- st_as_sf(neon_sites_t, coords = c("field_longitude", "field_latitude"), crs = 4326)


```

### Locate regions of interest

```{r}
# Validate (due to native crs issues), subset neon sites outside of regions. 
na_ecoregions_3_validated <- st_make_valid(na_ecoregions_3)
st_crs(na_ecoregions_3_validated)
neon_sites_in_regions <- neon_sites_sp[na_ecoregions_3_validated,]

# isolate ecoregions that contain Neon Sites
neon_sites_ecoregions <- na_ecoregions_3_validated[neon_sites_sp,]
```

#### Make some plots including the ecoregions & Neon Sites

```{r}
ggplot() + 
  geom_sf(na_ecoregions_3_validated, mapping = aes()) +
  geom_sf(neon_sites_sp, mapping = aes()) +
  theme_bw() + 
  ggtitle("Locations of NEON sites in relation to ecoregions")

ggplot() + 
  geom_sf(na_ecoregions_3_validated, mapping = aes()) +
  geom_sf(neon_sites_ecoregions, mapping = aes(fill = "red")) +
  geom_sf(neon_sites_sp, mapping = aes()) +
  theme_bw() + 
  ggtitle("Regions Containing NEON sites")

ggplot() + 
  geom_sf(na_ecoregions_3_validated, mapping = aes()) +
  geom_sf(neon_sites_ecoregions, mapping = aes(fill = "Region")) +
  geom_sf(neon_sites_in_regions, mapping = aes()) +
  theme_bw() + 
  ggtitle("Level 3 EPA North American Ecoregions Containing NEON sites") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill = guide_legend(title = "Regions of Interest"))
  
ggsave(filename = "/home/jt-miller/Pollen-Project/pollen-project-mapping/output-figures/NAecoregion-Neon-sites/ecoregion-overall-map-w-neon.png", plot = last_plot(), width = 20, height = 10, units = "cm")

```

### Extract Regions of Interest, Create Sampling Heatmaps

```{r}
regions_of_interest <- neon_sites_ecoregions$NA_L3NAME

region_i_df <- as.data.frame(regions_of_interest)

neon_sites_of_interest <- neon_sites_in_regions$field_site_name
neon_i_df <- as.data.frame(neon_sites_of_interest)
```

```{r}
#write.table(region_i_df, file = "/home/jt-miller/Pollen-Project/Pollen-DB/processed-data/shape-outputs/ecoregions/na-lvl3-ecoregions-w-neon-sites.txt", 
          # sep = "\t", row.names=FALSE, col.names=FALSE, quote=FALSE)

write.table(neon_i_df, file = "/home/jt-miller/Pollen-Project/pollen-project-mapping/processed-data/shape-outputs/neon-sites/neon_sites_in_lvl3NAregions.txt", 
          sep = "\t", row.names=FALSE, col.names=FALSE, quote=FALSE)
```

### Creating a ggplot with labeled ecoregions for clarity?
```{r}
# Create centroids: 
x <- st_centroid(neon_sites_ecoregions)
df <- x %>% 
  mutate(id = row_number()) %>% 
  mutate(long = unlist(map(x$geometry,1)),
           lat = unlist(map(x$geometry,2)))


separated_coord <- df %>%
    
ggplot() + 
  geom_sf(na_ecoregions_3_validated, mapping = aes()) +
  geom_sf(neon_sites_ecoregions, mapping = aes(fill = "Region")) +
  geom_text(df, mapping = aes(x = long, y = lat, label = id), size = 4 ) +
  theme_bw() + 
  ggtitle("Level 3 EPA North American Ecoregions Containing NEON sites") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill="none")
ggsave(filename = "/home/jt-miller/Pollen-Project/pollen-project-mapping/output-figures/NAecoregion-Neon-sites/ecoregion-overall-map-w-numbers.png", plot = last_plot(), width = 25, height = 25, units = "cm")
```

### Create Ecoregion orientation Maps 
```{r}
# Testing grounds 
setwd("/home/jt-miller/Pollen-Project/pollen-project-mapping/output-figures/NAecoregion-Neon-sites/")
for(i in 1:length(regions_of_interest)){
  region_to_subset <- regions_of_interest[[i]]
  subset_of_region <- subset(na_ecoregions_3_validated, na_ecoregions_3_validated$NA_L3NAME == region_to_subset)
  neon_site_subset <- neon_sites_sp[subset_of_region,]
  
  ggplot() + 
  geom_sf(na_ecoregions_3_validated, mapping = aes()) +
  geom_sf(subset_of_region, mapping = aes(fill = "Region")) +
  geom_sf(neon_site_subset, mapping = aes()) +
  theme_bw() + 
  ggtitle(paste0("Neon Sites with", "\n", regions_of_interest[[i]])) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill="none")

  
  ggsave(file=paste0("region_of_interest_plot", regions_of_interest[[i]],".png"), plot = last_plot(), width = 20, height = 10, units = "cm")
  
}


test <- regions_of_interest[[1]]
test2 <- subset(na_ecoregions_3_validated, na_ecoregions_3_validated$NA_L3NAME == test)
neon_sites_test <- neon_sites_sp[test2,]

ggplot() + 
  geom_sf(na_ecoregions_3_validated, mapping = aes()) +
  geom_sf(test2, mapping = aes(fill = "Region")) +
  geom_sf(neon_sites_test, mapping = aes()) +
  theme_bw() + 
  ggtitle(paste0("Neon Sites with", "\n", regions_of_interest[[1]])) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill = guide_legend(title = "Regions of Interest"))
```


